---
title: "TidyTuesday | Coffee Ratings"
author: "fth"
format: gfm
editor: visual
---

## Coffee Rating Dataset

Exploring an older TidyTuesday dataset with new tools.

setup chunk:
```{r, setup}
#| warning: false

library(dplyr)
library(tidyr)
library(corrr) 

library(ggplot2)
library(gt)            
library(patchwork)

library(monochromeR)
library(PrettyCols)
library(emo)
# emo::ji_find("hot_beverage")
emo::ji("hot_beverage") -> cup 
gt_grob <- function(gt_object, ...){
# https://github.com/rstudio/gt/issues/961  
  out_name <- file.path(
    tempfile(pattern = "file", tmpdir = tempdir(), fileext = ".png")
  )
  
  gtsave(gt_object, out_name, ...)
  
  in_png <- png::readPNG(out_name)
  
  on.exit(file.remove(out_name), add=TRUE)
  
  grid::rasterGrob(in_png)
  
}

```

get the dataset: 

```{r, get data}
# tidytuesday coffee ---------------------------
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

```

run some calculations: 
```{r, do calcs}
# average rating by country  --
coffee_ratings %>% 
  group_by(country_of_origin) %>%
  summarize(
    mean_total_cup_pts = mean(total_cup_points)
  ) %>% 
  arrange(desc(mean_total_cup_pts)) -> average_rating_by_country


# average rating subgroups
coffee_ratings %>% 
  group_by(species, variety, processing_method) %>%
  summarize(
    mean_total_cup_pts = mean(total_cup_points, na.rm = TRUE),
    mean_alt = mean(altitude_mean_meters, na.rm = TRUE)
  ) %>% 
  arrange(desc(mean_total_cup_pts)) -> average_rating_groups

# unique coffees -- 
coffee_ratings %>% 
  group_by(country_of_origin) %>% 
  count %>% 
  arrange(desc(n)) -> unique_coffees

# top rated with many coffees -- 
unique_coffees %>% 
  left_join(average_rating_by_country, 
            by = 'country_of_origin') %>% 
  ungroup() %>% 
  head(15) %>% # want top ranked origins with large variety
  arrange(desc(mean_total_cup_pts)) %>% 
  head(10) -> unique_coffees 
```

removing some unnecessary cols, & filtering on various row splices

```{r, wrangling}
# wrangling // filtering ----
coffee_filtered <- coffee_ratings %>% 
                   select(total_cup_points, species, country_of_origin, region, harvest_year,
                          variety:category_two_defects,
                          altitude_mean_meters) 

coffee_filtered %>% 
  filter(
    country_of_origin %in% unique_coffees$country_of_origin
    ) -> coffee_filtered

```


creating some colors and themes
```{r, theming}
brownie <- generate_palette("brown", modification = "go_both_ways", 
                                   n_colours = 20, 
                                   view_palette = TRUE, 
                                   view_labels = FALSE)

brown_theme <- theme(plot.background = element_rect(fill = brownie[8], 
                                                    colour = brownie[8]),
                     panel.grid = element_line(colour = brownie[5]),
                     panel.background = element_rect(fill = brownie[8], 
                                                     colour = brownie[8]),
                     text = element_text(colour = brownie[15]),
                     axis.text = element_text(colour = brownie[17]),
                     plot.title = element_text(colour = brownie[20], hjust = 0),
                     plot.subtitle = element_text(colour = brownie[18], hjust = 0)) + 
                theme(
                  axis.text =   element_text(size = 12),
                  plot.title =  element_text(size = 20),
                  plot.subtitle = element_text(size = 12),
                  axis.title =  element_text(size = 12),
                  legend.text = element_text(size = 12)
                )

```


# gt tables

```{r, gt countries}
# gt -----------------

# most unique and average rating
gt_countries <- unique_coffees %>% 
                # left_join(average_rating_by_country, 
                #           by = 'country_of_origin') %>% 
                # ungroup() %>% 
                # head(15) %>% # want countries with high variety (i.e. not 1 coffee:country)
                # arrange(desc(mean_total_cup_pts)) %>%
                # head(10) %>% # top ranks per high distinct counts
                gt() %>% 
                data_color(
                  columns = mean_total_cup_pts, 
                  colors = brownie
                ) %>% 
                fmt_number(
                  columns = mean_total_cup_pts
                ) %>% 
                tab_header(
                  title = paste0("Origin Samples & Mean Total ", cup, " Points ")
                ) %>% 
                cols_label(
                  country_of_origin  =   "Origin",
                  n =                    "Distinct Samples",
                  mean_total_cup_pts =   "Avg Rating"
                ) %>% 
               tab_options(table.background.color = brownie[5],
                           table_body.hlines.color = brownie[3],
                           column_labels.border.bottom.color = brownie[10],
                           heading.border.bottom.color = brownie[10])

as_raw_html(gt_countries) # https://github.com/rstudio/gt/issues/104
```

```{r, gt corrs}
# top corrs to cup points
gt_corrs <-
  coffee_filtered %>% 
  select(total_cup_points, aroma:cupper_points) %>% 
  correlate() %>%
  focus(total_cup_points) %>% 
  arrange(desc(total_cup_points)) %>% 
  gt() %>% 
  data_color(
    columns = total_cup_points, 
    colors = brownie
  ) %>% 
  fmt_number(
    columns = total_cup_points
  ) %>%
  tab_header(
    title = paste0(cup, " Points & Rating Correlations")
  ) %>% 
  cols_label(
    total_cup_points  =   "Total Cup Correlation",
    term =                "Other Ratings",
  ) %>% 
  tab_options(table.background.color = brownie[5],
              table_body.hlines.color = brownie[3],
              column_labels.border.bottom.color = brownie[10],
              heading.border.bottom.color = brownie[10])

as_raw_html(gt_corrs)
```

# gg section 


```{r, ggplot2}
# gg -------------------
top_countries <- c("Ethiopia", "Kenya", "Uganda", "Costa Rica", "Colombia", "El Salvador")

gg_elev <- 
coffee_filtered %>%
  filter(country_of_origin %in% top_countries) %>% 
  filter(altitude_mean_meters < 2500) %>% 
  filter(altitude_mean_meters > 600) %>% 
  filter(total_cup_points > 50) %>% 
  mutate(
    country_of_origin = forcats::fct_reorder(country_of_origin, total_cup_points, .fun='mean')
  ) %>% 
  ggplot(.) + 
  geom_point(aes(x = altitude_mean_meters, y = total_cup_points, 
                 color = country_of_origin), size = 3) + 
  geom_smooth(aes(x = altitude_mean_meters, y = total_cup_points), 
              color = brownie[15], 
              se = FALSE, linetype = "longdash", size = 1.2) + 
  labs(
    title = "Cup Points & Elevation",
    x = "Farm Mean Elevation (m)",
    y = paste0("Total ", cup, " Points")
  ) + 
  scale_color_pretty_d("Autumn") + 
  theme_minimal() + 
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
  ) + 
  brown_theme #+ 
    #theme(aspect.ratio = 1)

gg_jitter <- 
coffee_filtered %>%
  filter(country_of_origin %in% top_countries) %>% 
  mutate(
    country_of_origin = forcats::fct_reorder(country_of_origin, total_cup_points, .fun='mean')
  ) %>% 
  ggplot(.) + 
  geom_jitter(aes(y = country_of_origin, x = total_cup_points, 
                  color = country_of_origin),
              size = 3, alpha = .9) +
  labs(
    title = "Top Rated Origins",
    x = paste0("Total ", cup, " Points"),
    y = "Origin"
  ) + 
  scale_color_pretty_d("Autumn") + 
  theme_minimal() + 
  theme(
    legend.title = element_blank(),
    legend.position = "none"
  ) + 
  brown_theme #+ 
    #theme(aspect.ratio = 1)


# create ggplot of gt info, gt does not jive well with patchwork
gg_countries <- 
unique_coffees %>% 
  # left_join(average_rating_by_country, 
  #           by = 'country_of_origin') %>% 
  # ungroup() %>% 
  # head(15) %>% # want countries with high variety (i.e. not 1 coffee:country)
  # arrange(desc(mean_total_cup_pts)) %>%
  # head(10) %>% 
  ggplot() + 
  geom_text(aes(x = n, y = mean_total_cup_pts, 
                label = country_of_origin, color = country_of_origin), size = 5) + 
  labs(
    title = "Average Origin Rating & Total Coffees Rated",
    x = "Unique Coffees Rated",
    y = paste0("Mean Total ", cup, " Points")
  ) +
  scale_color_pretty_d("Autumn") + 
  theme_minimal() + 
  theme(
    legend.title = element_blank(),
    legend.position = "none"
  ) + 
  brown_theme #+ 
    #theme(aspect.ratio = 1)

gg_corrs <- 
  coffee_filtered %>% 
  select(total_cup_points, aroma:cupper_points) %>% 
  correlate() %>%
  focus(total_cup_points) %>% 
  arrange(desc(total_cup_points)) %>% 
  ggplot(.) + 
  geom_col(aes(x = forcats::fct_rev(forcats::fct_reorder(term, total_cup_points)), y = total_cup_points), 
           fill = brownie[15]) +
    labs(
      title = paste0("Correlations to Total ", cup ," Points"),
      x = "",
      y = ""
    ) + 
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    brown_theme #+
  #theme(aspect.ratio = 1)


```

## EDA

of the top rated origins, below shows the distributions of cup ratings for each origin using jitter plots. 
```{r}
gg_jitter
```


exploring the slight relationship seen when viewing cup rating as a function of farm elevation.
```{r}
gg_elev
```

correlations between total cup rating & various contributing factors/ratings 
```{r}
gg_corrs
```

country ratings among the most diverse origins
```{r}
gg_countries
```


# patchwork 

```{r, patchwork}

# patchwork -----------------------------------------------------------------------------
# layout <- "
# AC
# BD
# "
# gg_jitter + gg_elev + gg_corrs + gg_countries +
#   plot_layout(design = layout) + 
#   plot_annotation(
#     title = 'Coffee Rating & Origin Exploration',
#     subtitle = 'A Story of Total Cup Ratings & Other Features',
#     caption = 'TidyTuesday Coffee Rating Dataset  
#                Using some new friends: patchwork, monochromeR, emo & PrettyCols +   
#                some old faithfuls: dplyr, tidyr, corrr, & ggplot2',
#     theme = brown_theme + theme(plot.title = element_text(size = 26))
#   ) 

# wrap_plots(gg_jitter, gg_corrs, gg_elev, gg_countries, 
#            ncol = 2) + 
#   plot_annotation(
#     title = 'Coffee Rating & Origin Exploration',
#     subtitle = 'A Story of Total Cup Ratings & Other Features',
#     caption = 'TidyTuesday Coffee Rating Dataset
#                Using some new friends: patchwork, monochromeR, emo & PrettyCols +
#                some old faithfuls: dplyr, tidyr, corrr, & ggplot2',
#     theme = brown_theme + theme(plot.title = element_text(size = 26))
#   )


```

